---
title: "EDA lab"
author: "SID 460375939"
date: "28 August 2018"
output:
  html_document:
    df_print: paged
---

# Assignment 1

### Due date: 29/08/2018 - 11:59pm

### Worth: 5% 

Your assignment, if you choose to accept it (and you must), is to perform an exploratoray data analysis (EDA)
on the data used on the major project. The dataset for the major project can be found in the resources section
on Canvas in the file:

* "nutmstatData2018.csv"

For the first lab your task is to

1) Calculate summary statistics taking into account the variable definitions containted in the 
"nutmstatDataItems2018.csv" file. Be careful to treat categorical variables appropriately.

2) Take care of missing values! Determine the percentage
of missing values in each row/column. Summarise this information in some way.

3) Partition the dataset by taking subsets of rows or columns to use as many rows/columns as posssible.

So could do full values as a partition, and some partitions of some missing values.

4) Determine what variables might be used as predictors.

5) Plot some (not all historgrams). Only plot what you think is useful.

only continuous data.

6) Cluster the data. Determine a reasonable number of clusters.

7) Perform a principlan component analysis. Do the principal components capture 
potential categorial predictors in some way.

Students are to submit a .pdf file of their analysis.

Some code is given below to start you off. Some more code will become available in the coming days.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


#library(tidyverse)
```

```{r Read data, echo=FALSE}

dat_orig <- read.csv("nutmstatData2018.csv",header=TRUE)
dim(dat_orig)

dat <-  dat_orig

colNames <- colnames(dat_orig)

```

# Recode variables

```{r Recode variables, echo=FALSE}

# for (i in 1:length(colNames)) { cat("categoricalList[",i,"] <- FALSE # ",colNames[i],"\n"); }

categoricalList <- c()
categoricalList[ 1 ] <- FALSE #  BMISC 
categoricalList[ 2 ] <- FALSE #  AGEC 
categoricalList[ 3 ] <- TRUE #  SMSBC 
categoricalList[ 4 ] <- TRUE #  COBBC 
categoricalList[ 5 ] <- TRUE #  FEMLSBC 
categoricalList[ 6 ] <- FALSE #  PHDKGWBC 
categoricalList[ 7 ] <- FALSE #  PHDCMHBC 
categoricalList[ 8 ] <- FALSE #  EXLWTBC 
categoricalList[ 9 ] <- FALSE #  EXLWMBC 
categoricalList[ 10 ] <- FALSE #  EXLWVBC 
categoricalList[ 11 ] <- FALSE #  PHDCMWBC 
categoricalList[ 12 ] <- FALSE #  BMR 
categoricalList[ 13 ] <- FALSE #  EIBMR1 
categoricalList[ 14 ] <- TRUE #  SF2SA1QN 
categoricalList[ 15 ] <- TRUE #  INCDEC 
categoricalList[ 16 ] <- TRUE #  DIABBC 
categoricalList[ 17 ] <- TRUE #  HCHOLBC 
categoricalList[ 18 ] <- TRUE #  HSUGBC 
categoricalList[ 19 ] <- TRUE #  HYPBC 
categoricalList[ 20 ] <- FALSE #  ENERGYT1 
categoricalList[ 21 ] <- FALSE #  ENRGYT1 
categoricalList[ 22 ] <- FALSE #  MOISTT1 
categoricalList[ 23 ] <- FALSE #  PROTT1 
categoricalList[ 24 ] <- FALSE #  FATT1 
categoricalList[ 25 ] <- FALSE #  CHOWSAT1 
categoricalList[ 26 ] <- FALSE #  CHOWOAT1 
categoricalList[ 27 ] <- FALSE #  STARCHT1 
categoricalList[ 28 ] <- FALSE #  SUGART1 
categoricalList[ 29 ] <- FALSE #  FIBRET1 
categoricalList[ 30 ] <- FALSE #  ALCT1 
categoricalList[ 31 ] <- FALSE #  PREVAT1 
categoricalList[ 32 ] <- FALSE #  PROVAT1 
categoricalList[ 33 ] <- FALSE #  RETEQT1 
categoricalList[ 34 ] <- FALSE #  B1T1 
categoricalList[ 35 ] <- FALSE #  B2T1 
categoricalList[ 36 ] <- FALSE #  B3T1 
categoricalList[ 37 ] <- FALSE #  NIACINT1 
categoricalList[ 38 ] <- FALSE #  FOLATT1 
categoricalList[ 39 ] <- FALSE #  FOLICT1 
categoricalList[ 40 ] <- FALSE #  TOTFOLT1 
categoricalList[ 41 ] <- FALSE #  FOLEQT1 
categoricalList[ 42 ] <- FALSE #  B6T1 
categoricalList[ 43 ] <- FALSE #  B12T1 
categoricalList[ 44 ] <- FALSE #  VITCT1 
categoricalList[ 45 ] <- FALSE #  VITET1 
categoricalList[ 46 ] <- FALSE #  CALCT1 
categoricalList[ 47 ] <- FALSE #  IODINET1 
categoricalList[ 48 ] <- FALSE #  IRONT1 
categoricalList[ 49 ] <- FALSE #  MAGT1 
categoricalList[ 50 ] <- FALSE #  PHOST1 
categoricalList[ 51 ] <- FALSE #  POTAST1 
categoricalList[ 52 ] <- FALSE #  SELT1 
categoricalList[ 53 ] <- FALSE #  SODIUMT1 
categoricalList[ 54 ] <- FALSE #  ZINCT1 
categoricalList[ 55 ] <- FALSE #  CAFFT1 
categoricalList[ 56 ] <- FALSE #  CHOLT1 
categoricalList[ 57 ] <- FALSE #  SATFATT1 
categoricalList[ 58 ] <- FALSE #  MUFATT1 
categoricalList[ 59 ] <- FALSE #  PUFATT1 
categoricalList[ 60 ] <- FALSE #  LAT1 
categoricalList[ 61 ] <- FALSE #  ALAT1 
categoricalList[ 62 ] <- FALSE #  LCN3T1 
categoricalList[ 63 ] <- FALSE #  TRANST1 
categoricalList[ 64 ] <- FALSE #  PROPER1 
categoricalList[ 65 ] <- FALSE #  FATPER1 
categoricalList[ 66 ] <- FALSE #  LAPER1 
categoricalList[ 67 ] <- FALSE #  ALAPER1 
categoricalList[ 68 ] <- FALSE #  CHOPER1 
categoricalList[ 69 ] <- FALSE #  SUGPER1 
categoricalList[ 70 ] <- FALSE #  STARPER1 
categoricalList[ 71 ] <- FALSE #  ALCPER1 
categoricalList[ 72 ] <- FALSE #  SATPER1 
categoricalList[ 73 ] <- FALSE #  TRANPER1 
categoricalList[ 74 ] <- FALSE #  FIBRPER1 
categoricalList[ 75 ] <- FALSE #  MONOPER1 
categoricalList[ 76 ] <- FALSE #  POLYPER1 
categoricalList[ 77 ] <- FALSE #  ADTOTSE 
categoricalList[ 78 ] <- TRUE #  BDYMSQ04 
categoricalList[ 79 ] <- FALSE #  DIASTOL 
categoricalList[ 80 ] <- TRUE #  DIETQ12 
categoricalList[ 81 ] <- TRUE #  DIETQ14 
categoricalList[ 82 ] <- TRUE #  DIETQ5 
categoricalList[ 83 ] <- TRUE #  DIETQ8 
categoricalList[ 84 ] <- TRUE #  DIETRDI 
categoricalList[ 85 ] <- TRUE #  SABDYMS 
categoricalList[ 86 ] <- TRUE #  SEX 
categoricalList[ 87 ] <- FALSE #  SLPTIME 
categoricalList[ 88 ] <- TRUE #  SMKDAILY 
categoricalList[ 89 ] <- TRUE #  SMKSTAT 
categoricalList[ 90 ] <- FALSE #  SYSTOL 
categoricalList[ 91 ] <- TRUE #  ABSPID 
categoricalList[ 92 ] <- FALSE #  GRAINS1N 
categoricalList[ 93 ] <- FALSE #  WHOLGR1N 
categoricalList[ 94 ] <- FALSE #  REFGRA1N 
categoricalList[ 95 ] <- FALSE #  VEGLEG1N 
categoricalList[ 96 ] <- FALSE #  GREENS1N 
categoricalList[ 97 ] <- FALSE #  VGORSV1N 
categoricalList[ 98 ] <- FALSE #  STARCH1N 
categoricalList[ 99 ] <- FALSE #  LEGVEG1N 
categoricalList[ 100 ] <- FALSE #  OTHVEG1N 
categoricalList[ 101 ] <- FALSE #  FRUIT1N 
categoricalList[ 102 ] <- FALSE #  FRJUIC1N 
categoricalList[ 103 ] <- FALSE #  DAIRY1N 
categoricalList[ 104 ] <- FALSE #  MEAT1N 
categoricalList[ 105 ] <- FALSE #  RDMTL1N 
categoricalList[ 106 ] <- FALSE #  RDMTLU1N 
categoricalList[ 107 ] <- FALSE #  RDMTN1N 
categoricalList[ 108 ] <- FALSE #  RDMTNU1N 
categoricalList[ 109 ] <- FALSE #  RDMTNP1N 
categoricalList[ 110 ] <- FALSE #  PLTYL1N 
categoricalList[ 111 ] <- FALSE #  PLTYLU1N 
categoricalList[ 112 ] <- FALSE #  FISH1N 
categoricalList[ 113 ] <- FALSE #  EGGS1N 
categoricalList[ 114 ] <- FALSE #  LEGMT1N 
categoricalList[ 115 ] <- FALSE #  NUTS1N 
categoricalList[ 116 ] <- FALSE #  WATER1N 
categoricalList[ 117 ] <- FALSE #  UNSAT1N 
categoricalList[ 118 ] <- FALSE #  FRESUG1N 
categoricalList[ 119 ] <- FALSE #  ADDSUG1N 
categoricalList[ 120 ] <- FALSE #  WATERG1N 
categoricalList[ 121 ] <- FALSE #  PEFRESD1 
categoricalList[ 122 ] <- FALSE #  PEADDSD1 
categoricalList[ 123 ] <- TRUE #  ALTNTR 
categoricalList[ 124 ] <- TRUE #  ALTRESB 
categoricalList[ 125 ] <- TRUE #  APOBNTR 
categoricalList[ 126 ] <- TRUE #  APOBRESB 
categoricalList[ 127 ] <- TRUE #  B12RESB 
categoricalList[ 128 ] <- TRUE #  BIORESPC 
categoricalList[ 129 ] <- TRUE #  CHOLNTR 
categoricalList[ 130 ] <- TRUE #  CHOLRESB 
categoricalList[ 131 ] <- TRUE #  CVDMEDST 
categoricalList[ 132 ] <- TRUE #  DIAHBRSK 
categoricalList[ 133 ] <- TRUE #  FASTSTAD 
categoricalList[ 134 ] <- TRUE #  FOLATREB 
categoricalList[ 135 ] <- TRUE #  GGTNTR 
categoricalList[ 136 ] <- TRUE #  GGTRESB 
categoricalList[ 137 ] <- TRUE #  GLUCFPD 
categoricalList[ 138 ] <- TRUE #  GLUCFREB 
categoricalList[ 139 ] <- TRUE #  HBA1PREB 
categoricalList[ 140 ] <- TRUE #  HDLCHREB 
categoricalList[ 141 ] <- TRUE #  LDLNTR 
categoricalList[ 142 ] <- TRUE #  LDLRESB 
categoricalList[ 143 ] <- TRUE #  TRIGNTR 
categoricalList[ 144 ] <- TRUE #  TRIGRESB 

for (i in 1:length(categoricalList)) {
  if (categoricalList[ i ]) {
      dat[,i] <- as.factor(dat[  ,i])
  }
}




# for (i in 1:length(colNames)) { cat("excludeList[[",i,"]] <- c() # ",colNames[i],"\n"); }

excludeList <- list()

excludeList[[ 1 ]] <- c(0,98,99) #  BMISC 
excludeList[[ 2 ]] <- c() #  AGEC 
excludeList[[ 3 ]] <- c(0) #  SMSBC 
excludeList[[ 4 ]] <- c() #  COBBC 
excludeList[[ 5 ]] <- c() #  FEMLSBC 
excludeList[[ 6 ]] <- c(0,997,998,999) #  PHDKGWBC 
excludeList[[ 7 ]] <- c(0,998,999) #  PHDCMHBC 
excludeList[[ 8 ]] <- c(9996,9999) #  EXLWTBC 
excludeList[[ 9 ]] <- c(9996, 9999) #  EXLWMBC 
excludeList[[ 10 ]] <- c(9996 , 9999) #  EXLWVBC 
excludeList[[ 11 ]] <- c(0,998,999) #  PHDCMWBC 
excludeList[[ 12 ]] <- c(99998) #  BMR 
excludeList[[ 13 ]] <- c(998) #  EIBMR1 
excludeList[[ 14 ]] <- c(0,99) #  SF2SA1QN 
excludeList[[ 15 ]] <- c(0,98,99) #  INCDEC 
excludeList[[ 16 ]] <- c() #  DIABBC 
excludeList[[ 17 ]] <- c() #  HCHOLBC 
excludeList[[ 18 ]] <- c() #  HSUGBC 
excludeList[[ 19 ]] <- c() #  HYPBC 
excludeList[[ 20 ]] <- c() #  ENERGYT1 
excludeList[[ 21 ]] <- c() #  ENRGYT1 
excludeList[[ 22 ]] <- c() #  MOISTT1 
excludeList[[ 23 ]] <- c() #  PROTT1 
excludeList[[ 24 ]] <- c() #  FATT1 
excludeList[[ 25 ]] <- c() #  CHOWSAT1 
excludeList[[ 26 ]] <- c() #  CHOWOAT1 
excludeList[[ 27 ]] <- c() #  STARCHT1 
excludeList[[ 28 ]] <- c() #  SUGART1 
excludeList[[ 29 ]] <- c() #  FIBRET1 
excludeList[[ 30 ]] <- c() #  ALCT1 
excludeList[[ 31 ]] <- c() #  PREVAT1 
excludeList[[ 32 ]] <- c() #  PROVAT1 
excludeList[[ 33 ]] <- c() #  RETEQT1 
excludeList[[ 34 ]] <- c() #  B1T1 
excludeList[[ 35 ]] <- c() #  B2T1 
excludeList[[ 36 ]] <- c() #  B3T1 
excludeList[[ 37 ]] <- c() #  NIACINT1 
excludeList[[ 38 ]] <- c() #  FOLATT1 
excludeList[[ 39 ]] <- c() #  FOLICT1 
excludeList[[ 40 ]] <- c() #  TOTFOLT1 
excludeList[[ 41 ]] <- c() #  FOLEQT1 
excludeList[[ 42 ]] <- c() #  B6T1 
excludeList[[ 43 ]] <- c() #  B12T1 
excludeList[[ 44 ]] <- c() #  VITCT1 
excludeList[[ 45 ]] <- c() #  VITET1 
excludeList[[ 46 ]] <- c() #  CALCT1 
excludeList[[ 47 ]] <- c() #  IODINET1 
excludeList[[ 48 ]] <- c() #  IRONT1 
excludeList[[ 49 ]] <- c() #  MAGT1 
excludeList[[ 50 ]] <- c() #  PHOST1 
excludeList[[ 51 ]] <- c() #  POTAST1 
excludeList[[ 52 ]] <- c() #  SELT1 
excludeList[[ 53 ]] <- c() #  SODIUMT1 
excludeList[[ 54 ]] <- c() #  ZINCT1 
excludeList[[ 55 ]] <- c() #  CAFFT1 
excludeList[[ 56 ]] <- c() #  CHOLT1 
excludeList[[ 57 ]] <- c() #  SATFATT1 
excludeList[[ 58 ]] <- c() #  MUFATT1 
excludeList[[ 59 ]] <- c() #  PUFATT1 
excludeList[[ 60 ]] <- c() #  LAT1 
excludeList[[ 61 ]] <- c() #  ALAT1 
excludeList[[ 62 ]] <- c() #  LCN3T1 
excludeList[[ 63 ]] <- c() #  TRANST1 
excludeList[[ 64 ]] <- c() #  PROPER1 
excludeList[[ 65 ]] <- c() #  FATPER1 
excludeList[[ 66 ]] <- c() #  LAPER1 
excludeList[[ 67 ]] <- c() #  ALAPER1 
excludeList[[ 68 ]] <- c() #  CHOPER1 
excludeList[[ 69 ]] <- c() #  SUGPER1 
excludeList[[ 70 ]] <- c() #  STARPER1 
excludeList[[ 71 ]] <- c() #  ALCPER1 
excludeList[[ 72 ]] <- c() #  SATPER1 
excludeList[[ 73 ]] <- c() #  TRANPER1 
excludeList[[ 74 ]] <- c() #  FIBRPER1 
excludeList[[ 75 ]] <- c() #  MONOPER1 
excludeList[[ 76 ]] <- c() #  POLYPER1 
excludeList[[ 77 ]] <- c(9996,9999) #  ADTOTSE 
excludeList[[ 78 ]] <- c(0,6) #  BDYMSQ04 
excludeList[[ 79 ]] <- c(0,998,999) #  DIASTOL 
excludeList[[ 80 ]] <- c(0,6) #  DIETQ12 
excludeList[[ 81 ]] <- c(0,6) #  DIETQ14 
excludeList[[ 82 ]] <- c(0) #  DIETQ5 
excludeList[[ 83 ]] <- c(0) #  DIETQ8 
excludeList[[ 84 ]] <- c(0,3) #  DIETRDI 
excludeList[[ 85 ]] <- c(0,8,9) #  SABDYMS 
excludeList[[ 86 ]] <- c() #  SEX 
excludeList[[ 87 ]] <- c(9998,9999) #  SLPTIME 
excludeList[[ 88 ]] <- c() #  SMKDAILY 
excludeList[[ 89 ]] <- c(0) #  SMKSTAT 
excludeList[[ 90 ]] <- c(0,998,999) #  SYSTOL 
excludeList[[ 91 ]] <- c() #  ABSPID 
excludeList[[ 92 ]] <- c() #  GRAINS1N 
excludeList[[ 93 ]] <- c() #  WHOLGR1N 
excludeList[[ 94 ]] <- c() #  REFGRA1N 
excludeList[[ 95 ]] <- c() #  VEGLEG1N 
excludeList[[ 96 ]] <- c() #  GREENS1N 
excludeList[[ 97 ]] <- c() #  VGORSV1N 
excludeList[[ 98 ]] <- c() #  STARCH1N 
excludeList[[ 99 ]] <- c() #  LEGVEG1N 
excludeList[[ 100 ]] <- c() #  OTHVEG1N 
excludeList[[ 101 ]] <- c() #  FRUIT1N 
excludeList[[ 102 ]] <- c() #  FRJUIC1N 
excludeList[[ 103 ]] <- c() #  DAIRY1N 
excludeList[[ 104 ]] <- c() #  MEAT1N 
excludeList[[ 105 ]] <- c() #  RDMTL1N 
excludeList[[ 106 ]] <- c() #  RDMTLU1N 
excludeList[[ 107 ]] <- c() #  RDMTN1N 
excludeList[[ 108 ]] <- c() #  RDMTNU1N 
excludeList[[ 109 ]] <- c() #  RDMTNP1N 
excludeList[[ 110 ]] <- c() #  PLTYL1N 
excludeList[[ 111 ]] <- c() #  PLTYLU1N 
excludeList[[ 112 ]] <- c() #  FISH1N 
excludeList[[ 113 ]] <- c() #  EGGS1N 
excludeList[[ 114 ]] <- c() #  LEGMT1N 
excludeList[[ 115 ]] <- c() #  NUTS1N 
excludeList[[ 116 ]] <- c() #  WATER1N 
excludeList[[ 117 ]] <- c() #  UNSAT1N 
excludeList[[ 118 ]] <- c() #  FRESUG1N 
excludeList[[ 119 ]] <- c() #  ADDSUG1N 
excludeList[[ 120 ]] <- c() #  WATERG1N 
excludeList[[ 121 ]] <- c() #  PEFRESD1 
excludeList[[ 122 ]] <- c() #  PEADDSD1 
excludeList[[ 123 ]] <- c(0,8) #  ALTNTR 
excludeList[[ 124 ]] <- c(97,98) #  ALTRESB 
excludeList[[ 125 ]] <- c(0,8) #  APOBNTR 
excludeList[[ 126 ]] <- c(97,98) #  APOBRESB 
excludeList[[ 127 ]] <- c(97,98) #  B12RESB 
excludeList[[ 128 ]] <- c(0) #  BIORESPC 
excludeList[[ 129 ]] <- c(0,8) #  CHOLNTR 
excludeList[[ 130 ]] <- c(97,98) #  CHOLRESB 
excludeList[[ 131 ]] <- c(0,8) #  CVDMEDST 
excludeList[[ 132 ]] <- c(0,8) #  DIAHBRSK 
excludeList[[ 133 ]] <- c(0) #  FASTSTAD 
excludeList[[ 134 ]] <- c(97,98) #  FOLATREB 
excludeList[[ 135 ]] <- c(8) #  GGTNTR 
excludeList[[ 136 ]] <- c(97,98) #  GGTRESB 
excludeList[[ 137 ]] <- c(0,8) #  GLUCFPD 
excludeList[[ 138 ]] <- c(97,98) #  GLUCFREB 
excludeList[[ 139 ]] <- c(7,8) #  HBA1PREB 
excludeList[[ 140 ]] <- c(7,8) #  HDLCHREB 
excludeList[[ 141 ]] <- c(0,8) #  LDLNTR 
excludeList[[ 142 ]] <- c(97,98) #  LDLRESB 
excludeList[[ 143 ]] <- c(0,8) #  TRIGNTR 
excludeList[[ 144 ]] <- c(97,98) #  TRIGRESB 

for (i in 1:length(excludeList)) {
  dat[ dat[,i]%in%excludeList[[ i ]],i] <- NA
}

``` 


# Plot continous variables

```{r Plot continous variables, echo=FALSE}

for (i in 1:length(categoricalList)) {
  if (!categoricalList[[ i ]]) {
       plot(density(dat[,i],na.rm=TRUE),main=colNames[i])
  }
}

``` 

## Plot Categorical Data

```{r Plot categorical variables}
for (i in 1:length(categoricalList)) {
  if (categoricalList[[ i ]]) {
       plot(dat[,i],main=colNames[i])
  }
}
```

# Summarise continous variables

```{r Summarise continous variables, echo=FALSE}


n = nrow(dat)
p = ncol(dat)

nCont = sum(!categoricalList)
nCat  = sum(categoricalList)

indsCont = which(!categoricalList)
indsCat  = which(categoricalList)

summaryTable = matrix(0,nCont,7)
for (i in 1:nCont) {
  sumVec = summary(dat[,indsCont[i]])
  summaryTable[i,1:length(sumVec)] = sumVec 
}
summaryTable[,7] = 100*summaryTable[,7]/n

summaryTable = signif(summaryTable,4)



 
colnames(summaryTable) = c("Min.","1st Qu.","Median","Mean","3rd Qu.","Max.","Perc NAs")

library(huxtable)
ht = hux( summaryTable ,  add_colnames = TRUE)
ht$Variable = c("Variable",colNames[indsCont])
 
prettyNumFun <-  list(function(x) formatC(x, 2, format = "f", width = 11, big.mark = ","))
number_format(ht)[-1,1:7] = prettyNumFun

#right_padding(ht) <- 20
#left_padding(ht)  <- 20

#bold(ht)[1, ]          <- TRUE
#bottom_border(ht)[1, ] <- 1



caption(ht) <- 'Summary Statistics of Continuous variables'

#col_width(ht) <- 15
ht <- set_align(ht,1:(nCont+1), 1:7, 'right')
 
ht

```
 
## Summarise Categorical Variables

Removing "ABSDIB" since the data recorded for that is not properly formatted, so we will just assume the NA percentage is 100%. Also recording the percentages of the table values rather than the count.

```{r Summarise categorical Variables}
remove = NULL
for(i in 1:nCat) {
  if(i == 20) {
      next
    }
  print(colNames[indsCat[i]])
  print((signif(100*(table(dat[,indsCat[i]])/n),4)))
  print("Perc of NA")
  print((n - sum(table(dat[indsCat[i]])))/n)
}

```

## Partitioning the Dataset

So looking at the data, we want to partition the dataset so we have a complete data set or a data set without any missing values. We will remove all the rows with missing data with any of the continuous data (specifically the not applicable data), then we will remove the categorical data with any missing values. Also removing the variable "ABSDIB" is also important as it was not properly formatted.

```{r Partitioning a full dataset}
compdat <- dat

removerows = NULL
ncomp = nrow(compdat)
for(i in 1:ncomp) {
  removerows[i] <- sum(is.na(compdat[i,indsCont]))/ncomp
}

deleterow = NULL
j=0
for(i in 1:ncomp) {
  if(removerows[i] > 0) {
    j <- j + 1
    deleterow[j] <- i
  }
}
compdat <- compdat[-deleterow,]

newN = nrow(compdat)
remove = NULL
for(i in 1:nCat) {
  if(i == 20) {
    remove = c(remove, i)
  }
  if((newN - sum(table(compdat[indsCat[i]])))/newN > 0) {
    remove = c(remove,i)
  } 
}
compdat[,indsCat[remove]] <- list(NULL)

dim(compdat)
```

## Variables for Responses

Some good response variables could be "BMISC" (BMI), "PHDKGWBC" (Weight), "PHDCMWBC" (Waist Size) and "DIABBC" (Diabetes, categorical), just going by the data dictionary and powerpoint slide.

## Histograms with Complete Data (No NAs)

Plotting the histograms for the most likely response variables, as well as some interesting predictor variables. Here are some histograms for the response variable.
```{r Histogram of possible responses}
hist(compdat$BMISC, main = "BMI", xlab = "Body Mass Index")
hist(compdat$PHDKGWBC, main = "Weight", xlab = "Kg")
hist(compdat$PHDCMWBC, main = "Waist Size", xlab = "Centimetres")
```
For the predictors, "PHDCMHBC" (Height) would be a useful predictor as it is highly correlated to weight naturally, which could mean that if there are heavier, then they taller. Age is another important factor to consider as well.
```{r Histogram of possible predictors}
hist(compdat$PHDCMHBC, main = "Height", xlab = "Centimetres")
hist(compdat$AGEC, main = "Age", xlab = "in Years")
```

## k-Means Clustering
First determining the amount of clusters that should be used, (ripped from "https://www.statmethods.net/advstats/cluster.html", to aid in analysis), and then plotting the clusters as such.
```{r Cluster}
wss <- (nrow(compdat)-1)*sum(apply(compdat,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(compdat, 
  	centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")

clusters <- kmeans(compdat, 12)
str(clusters)
library(cluster)
clusplot(compdat, clusters$cluster, color=TRUE, shade=TRUE, 
  	labels=2, lines=0)

```

## Principal Component Analysis
First changing the categorical data to numerical, then following with the principal component analysis.
```{r PCA}
for (i in 1:dim(compdat)[2]) {
    compdat[,i] <- as.numeric(compdat[  ,i])
}
pca = prcomp(compdat)
str(pca)
dim(pca$x)
```

As you can see, there are too many variables, thus it does not capture any single variable as of now.